HSC Synthetic Pipeline I: Basic Design and Characterization of the Photometric Performance of HSC

### 2017-04-12 draft/ms_v1.tex
CModel: magnitude

@@

    The quality of the DR1 data have been vetted in \citet{HSCDR1} via cross-checking
    with the Panoramic Survey Telescope and Rapid Response System (Pan-STARRS) 1
    imaging survey (\citealt{Schlafly2012, Tonry2012, Magnier2013}), examining the
    behaviors of the stellar sequence and the shape of the PSF models.
    Generally speaking, the PSF photometry is accurate to 1--2\% level, and the
    astrometry is accurate to ${\sim}10$ and 40 max internally and externally.

@@

\subsubsection{Input magnitude and the $\mathrm{S}/\mathrm{N}$ of PSF photometry}

@@

    The overall performance of PSF photometry is excellent.
    In $i$-band, the median magnitude difference for \tract{} with 0.5\asec{} seeing is
    around ${\sim}0.015$ mag (${\sim}1.5$\% accuracy) at
    $i_{\mathrm{Input}}{\sim}19.0$ mag.
    At $i_{\mathrm{Input}}{\sim}24.0$ mag, the average accuracy of PSF
    magnitude keeps at ${\sim}3$\% level (${\sim}0.036$ mag average difference).
    Down to $i_{\mathrm{Input}}{\sim}25.2$ mag, the average accuracy decreases to
    ${\sim}7$\% level with an average ${\sim}0.072$ mag difference.
    The \tract{} with worse seeing shows similar accuracy of PSF photometry at
    $i_{\mathrm{Input}}<23.5$ mag.
    At faint end, the accuracy becomes slightly worse: ${\sim}5$\% accuracy around
    $i_{\mathrm{Input}}{\sim}24.0$ mag; and ${\sim}11$\% accuracy around
    $i_{\mathrm{Input}}{\sim}25.2$ mag.
    \citet{HSCDR1} evaluates the accuracy of PSF photometry via external comparisons
    with the PS1 PV2 and SDSS data at $i<21$ mag, and finds a 1-2\% level accuracy,
    which is consistent with our results.
    At fainter magnitude, external comparisons become difficult due to the lack of deep
    imaging data at HSC level and the impact of the color-term when comparing photometry
    using different filters.
    The results from \synpipe{} hence provide more useful evaluations down to the
    detection limit\footnote{Strictly speaking, the uncertainties here should be
    considered as lower limit as the \synpipe{} now does not consider the uncertainties
    in PSF modeling.}.

@@

        \item In Fig \ref{fig:cmodel_mag}, we evaluate the accuracies of the \forced{}
            \cmodel{} magnitudes in all five bands.

        \item \plan{Generally speaking, the \forced{} \cmodel{} provides good estimates
            of magnitudes for galaxies between 20.0 and 25.0 magnitude in all filters and
            different seeing conditions. }

        \item \plan{For $i$-band the absolute flux accuracy is around 7\% level at
            $i{\sim}21.0$ mag, and it increases to ${\sim}10$\% at $i{\sim}25.0$ mag.
            The \forced{} \cmodel{} photometry slightly underestimate the total flux of
            galaxies at $21.0 < i < 23.0$ mag.
            The average magnitude difference is at 0.04 mag level.
            For $i > 23.0$ mag, the \forced{} \cmodel{} photometry is unbiased.}

        \item \plan{The \forced{} \cmodel{} magnitude error from \hscpipe{} only considers
            the per pixel noise, but not the systematic uncertainties involved in the
            modeling fitting process.
            Therefore it is no surprise that the typical uncertainty from the \synpipe{}
            test is larger than the magnitude error.}

        \item \plan{The accuracies of the \forced{} \cmodel{} magnitudes in other bands
            are similar to the $i$-band.
            The only exception is for galaxies with $y<24.0$ mag where \forced{} \cmodel{}
            seems to overestimate the flux slightly.}

        \item \plan{For galaxies with $i < 20.0$, the statistics are uncertain due to small
            sample size}

        \item \plan{Unlike the synthetic stars, the highly blended synthetic galaxies only
            have slightly biased photometry, where \forced{} \cmodel{} underestimates the
            fluxes of these galaxies by ${\sim}0.05-0.10$ mag.
            High blendedness does lead to larger scatter in photometric uncertainties.}

       \item \plan{We also test the \unforced{} \cmodel{} photometry in all five bands,
           and the results suggest similar accuracies.
           We highlight the differences between \forced{} and
           \unforced{} \cmodel{} photometry in Fig \ref{fig:cmodel_diff}.}

>>>>

\newcommand{\plus}{\raisebox{.4\height}{\scalebox{.6}{+}}}
\newcommand{\minus}{\raisebox{.4\height}{\scalebox{.8}{-}}}

@@

    The quality of the DR1 data have been vetted in \citet{HSCDR1} via cross-checking
    with the Panoramic Survey Telescope and Rapid Response System (Pan-STARRS) 1
    imaging survey (\citealt{Schlafly2012, Tonry2012, Magnier2013}), examining the
    behaviors of the stellar sequence and the shape of the PSF models.
    Generally speaking, the PSF photometry is accurate to 1--2\% level, and the
    astrometry is accurate to ${\sim}10$ and 40 mas internally and externally.

@@

\subsubsection{Input magnitude and the \s2n{} of PSF photometry}

@@

    The overall performance of PSF photometry is excellent.
    In $i$-band, the median magnitude difference for \tract{} with 0.45\asec{} seeing is
    around ${\sim}0.015$ mag (${\sim}1.5$\% accuracy) at
    $i_{\mathrm{Input}}{\sim}19.0$ mag.
    At $i_{\mathrm{Input}}{\sim}24.0$ mag, the average accuracy of PSF
    magnitude keeps at ${\sim}3$\% level (${\sim}0.036$ mag average difference).
    Down to $i_{\mathrm{Input}}{\sim}25.2$ mag, the average accuracy decreases to
    ${\sim}7$\% level with an average ${\sim}0.072$ mag difference.
    The \tract{} with worse seeing shows similar accuracy of PSF photometry at
    $i_{\mathrm{Input}}<23.5$ mag.
    At faint end, the accuracy becomes slightly worse: ${\sim}5$\% accuracy around
    $i_{\mathrm{Input}}{\sim}24.0$ mag; and ${\sim}11$\% accuracy around
    $i_{\mathrm{Input}}{\sim}25.2$ mag.
    \citet{HSCDR1} evaluates the accuracy of PSF photometry via external comparisons
    with the PS1 PV2 and SDSS data at $i<21$ mag, and finds a 1-2\% level accuracy,
    which is consistent with our results.
    At fainter magnitude, external comparisons become difficult due to the lack of deep
    imaging data at HSC level and the impact of the color-term when comparing photometry
    using different filters.
    The results from \synpipe{} hence provide more useful evaluations down to the
    detection limit\footnote{Strictly speaking, the uncertainties here should be
    considered as lower limit as the \synpipe{} now does not consider the uncertainties
    in PSF modeling.}.

@@

    In Fig \ref{fig:cmodel_mag}, we illustrate the photometric uncertainties of the 
    \cmodel{} magnitudes in each band using the same format with 
    Fig \ref{fig:psf_mag}.

    The overall performance of \cmodel{} photometry is very reasonable all the way 
    down to $i_{\mathrm{Input}}=25.2$ mag. 
    Comparing to the qualities of PSF photometry for stars, the magnitude differences
    of \cmodel{} show much larger scatter. 
    Given the diversity of properties of these galaxies and the complexities in the 
    model-fitting process, this difference is expected.  
    At the same time, we can see that the \cmodel{} algorithm in \hscpipe{} can 
    provide unbiased and consistent photometry for galaxies in different bands and 
    seeing conditions.

    In $i$-band, the median flux uncertainties of \forced{} \cmodel{} is ${\sim}7$\% 
    in the good-seeing \tract{} at $i_{\mathrm{Input}}{\sim}20.0$ mag, 
    and ${\sim}9$\% at $i_{\mathrm{Input}}{\sim}24.0$ mag.  
    Down to $i_{\mathrm{Input}}{\sim}25.0$ mag, the median uncertainty maintains 
    at ${\sim}11$\% level. 
    For the bad-seeing \tract{}, the performance is very similar in $i$-band. 
    From $i_{\mathrm{Input}}{\sim}20.0$ to $25.0$ mag, the median magnitude difference
    between output and input values changes from \plus{}$0.041$ mag to 
    \minus{}$0.005$ mag, which suggests that \forced{} \cmodel{} photometry is 
    unbiased down to very faint end. 
    It is worth noting that the photometric uncertainties estimated by \synpipe{}
    are significantly larger than the fiducial error from \hscpipe{} that does not 
    take into account the systematically uncertainties involved in the modeling 
    fitting process. 
    Although the synthetic galaxies here are described by the simple single-\ser{} 
    model, \cmodel{} algorithm can only approximate them to certain accuracy, 
    especially when the noise level, deblending uncertainty, and priors of model 
    parameters are considered. 
    Given that the real galaxies are more complex in structural details, these 
    uncertainties should be treated as lower limits.

    The performance of \forced{} \cmodel{} photometry is very consistent across all
    filters. 
    For $g$, $r$, and $z$-band, the photometric uncertainties are very similar to 
    the ones for $i$-band between $20.0$ to $25.0$ mag. 
    The \forced{} \cmodel{} magnitudes are also unbiased for these 3 filters 
    across the entire input magnitude range. 
    The uncertainties for $y$-band are slightly large, range from 9-17\% in the 
    same magnitude bins. 
    In addition, \forced{} \cmodel{} tends to overestimate the total flux in $y$-band
    at $y_{\mathrm{Input}}>24.0$ mag. 
    Mean magnitude differences change from \minus{}0.06 at 24.0 mag to \minus{}0.22
    mag at 25.0 mag.  
    Both worse seeing condition and higher background noise level can make it more 
    difficult to detect faint object on $y$-band image.
    At $y_{\mathrm{Input}}{\sim}24.0$, the average \s2n{} for \cmodel{} in $y$-band
    is already close to the detection threshold.

    As mentioned, the statistics for brighter galaxies are very uncertain due to 
    small sample size. 
    However, we do notice that, at $i_{\mathrm{Input}}<20.0$, the \forced{} 
    \cmodel{} photometry shows larger uncertainty and starts to systematically 
    underestimate the total flux.  
    This problem is confirmed via comparing with more carefully derived total fluxes 
    for bright galaxies (e.g. \citealt{HSCDR1}, Huang\etal in prep.). 
    It relates to the known issues of \hscpipe{} (e.g. over-deblend around bright 
    objects; inappropriate weighting of priors in \cmodel{}), but it may also suggest
    that, at the imaging depth of HSC, assumptions in \cmodel{} are no longer the 
    best choice in modeling these galaxies (e.g. for massive elliptical galaxies; 
    see Huang\etal in prep.~for more details).

    For the highly blended synthetic galaxies, we also see systematically 
    underestimated total flux (average offset ${\sim}0.1$ mag) and larger scatter 
    of photometric uncertainty in all five bands.  
    Meanwhile, the contrast between relative isolated and highly blended galaxies 
    in photometric performance is no longer as staggering as the one for synthetic
    stars (Fig \ref{fig:psf_mag}).

    We also test the \unforced{} \cmodel{} photometry in all five bands, and the 
    results suggest similar accuracies.
    We highlight the differences between \forced{} and \unforced{} \cmodel{} 
    photometry in $g$- and $y$-band in Fig \ref{fig:cmodel_diff}.
    The only noticeable difference is for galaxies brighter than $24.0$ mag in 
    $g$-band, their \forced{} \cmodel{} magnitudes are systematically brighter than 
    the \unforced{} ones.
